### 2024-10-09 word/example.qmd
added blank template

---
title: "Title of the example word file"
author:
  - Paul Raftery ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to p.raftery@berkeley.edu], Author2AtSameOrg ^1^, Author3atDifferentOrg ^[DifferentOrg]
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    reference_docx: "./template.docx" # this is the template containing the formatting for the Word file (font of headings, etc.)
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib # this points to a file containing your references for this paper
csl : elsevier-with-titles.csl # you can change citation style by changing this to refer to a different CSL file style
---

```{r knitrsetup, include = FALSE, cache = FALSE}
# knitr setup
# these are the defaults for how knitr reads and processes this file
# you can override them for specific chunks in the qmd file.
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  dev = "jpeg",
  cache = T,
  dpi = 300,
  fig.path = "figs",
  fig.show = "hold",
  fig.pos = "b",
  fig.width = 6.5, fig.height = 8.5, # these are the default figure sizes
  fig.path = "./word"
)

# str(knitr::opts_chunk$get()) # run this in the console for all knitr config options
```

```{r librariesandthemesettings, include = FALSE, cache = FALSE}
# load all required libraries and configure themes

require(pacman)

# load packages using pacman
pacman::p_load(
  tidyverse, lubridate, here, rmarkdown, scales, patchwork, magrittr, janitor, # general, commonly used
  styler, # auto style your code with this addin
  # ggpubr, # ggplot extension focuses on annotations and fitted models
  # ggExtra, # ggplot extension handy for things like marginal density curves on figures
  # sjstats, pwr, # anova results
  # slider, # moving averages
  # rstatix, #pipe friendly stats
  palmerpenguins # quarto functionality
)

options(
  scipen = 999, # turn off scientific notation,
  digits = 2, # set default significant digits to 2, note this only works after a decimal point, use round() or format() to reduce precision of larger numbers
  timeout = 60
)

# set directory
here::i_am("example.qmd")

# ## for example, theme minimal
# # set default theme for ggplot
# theme_set(theme_minimal())
#
# # modify as preferred, e.g.
# theme_update(plot.title = element_text(size = 14, colour = "grey20", face = "bold", hjust = 0.5),
#              plot.subtitle = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5, margin = margin(b = 10)),
#              plot.caption = element_text(size = 10, colour = "grey20", face = "italic", hjust = 0.5),
#              plot.background = element_rect(fill = "white", colour = NA),
#              panel.grid.minor = element_blank(),
#              panel.grid.major = element_blank(),
#              axis.text = element_text(size = 10),
#              strip.text = element_text(size = 10, color = "grey20", face = "bold"),
#              strip.background = element_blank())

# or theme light, with some  Cal colors
# figure theme setup
# set base font size for plots
theme_set(theme_light(base_size = 10))
# cleaner display of 'strips' above facets in plots
theme_update(
  strip.background = element_rect(fill = "#F5F5F5", color = "grey"),
  strip.text = element_text(color = "black"),
  strip.text.y = element_text(angle = 90)
)
#  lighten gridlines so just visible
theme_update(panel.grid = element_line(linetype = "dotted"))
# define a reasonably print and color-blind friendly color palette
# these also conveniently happen to be colors from the official UC Berkeley color palette. https://brand.berkeley.edu/identity-and-visual-design/color
cal_palette2 <- c("#3B7EA1", "#C4820E")
# with slight mod to dark yellow for better grayscale performance for 4 color scale
cal_palette4 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515")
# adding 5th color
cal_palette5 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515", "#D9661F")
```

```{r functions}
# user defined functions specific to this analysis
```

```{r dataprep, include = FALSE}
# Read in data, misc initial prep/cleaning tasks
# for example, list all the files in a folder that match a specific pattern (e.g. csv files)

files <- list.files(path = "./input_file_folder/", pattern = "*.csv")

# create a short function to read in those files one by one
# modify those files after you read them in,
# and return a dataframe
read_and_append_file <- function(fle) {
  df <- read_csv(str_c("./input_file_folder/", fle))
  # do whatever individual file level modifications you need here
  # for example, extract some string from the filename
  sss <- str_extract(string = fle, pattern = "(?<=df_).*(?=_USA_Clima)")
  df %<>% mutate(sourcefilename = fle, stringofinterest = sss)
  return(df)
}

# apply that function and return an assembled dataframe
df <- files %>% map_dfr(read_and_append_file)

# apply functions to tidy up the dataframe as necessary
df %<>% clean_names()
```

# Abstract {.unnumbered}
The above heading doesn't count towards heading numbers. This is handy for abstracts, nomenclature, etc. If you want to change the style and formatting of any of the headings in the finished document you can do so by modifying the Styles in the word template referenced above, 'template.docx'.

# First level heading
This is how you cite a reference [@rafteryQuantifyingEnergyLosses2018]. It must be in the bibliography file 'references.bib', which is in the bibtex file format and can be automatically output by a citation management software.

## Second level heading
You can and should perform any short calculations directly in the text. For example, there are `r nrow(iris)` rows in the Iris dataset. The mean  species length is `r mean(iris$Sepal.Length)`. Anything more substantial should be calculated (and well commented) in a separate code chunk. Note the number of significant digits is set globally above but can be overwritten as needed, for example that same value to more precision is `r format(mean(iris$Sepal.Length),digits=5)`. You can also reduce precision in large values above the decimal where needed, for example, report `r 10000*mean(iris$Sepal.Length)` as `r round(10000*mean(iris$Sepal.Length),-3)`.

```{r}
# you'll often need to perform calculations that require multiple lines of code and should be commented
# you can do these in a separate code chunk like this one
# any variables you define here can be used in any subsequent point in the doc
# that applies both to other chunks

weird_iris <- iris # create a version of the iris dataset with some random noise added to one of the results columns
weird_iris %<>%
  mutate(Sepal.Length = Sepal.Length * rnorm(nrow(weird_iris), mean = 1, sd = 0.1)) # add a multiplier of random noise of 0.1 standard deviation
```

The mean species length of the modified iris dataset is `r mean(weird_iris$Sepal.Length)`.

### Third level heading
Here's how you cross reference a figure \@ref(fig:yourfignamehere), or a table \@ref(tab:yourtablenamehere) or a section. This is how you include a footnote[^1].

[^1]: Your footnote text goes here. It can go anywhere in the .qmd file but it seems easiest to put it at the end of the paragraph where the footnote is first included.

```{r yourfignamehere, fig.cap="Your fig caption here", fig.width=6.5, fig.height=3}
# note that your fig name should ONLY USE letters OR NUMBERS as there are exceptional hard to debug issues that happen when you use symbols
#
# note that you should define the width and height in inches of your figure to fit and be legible on the template page type
# default is letter size, which means a max width of 6.5 inches and a max length of ~8.5inches depending on caption size
# another option is a single column size, which is a width of ~ 3.5 inches
p1 <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point()

p2 <- ggplot(weird_iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point()

# you can directly output a plot or multiple plots from this code chunk. however,
# often if you have more than one it's best to decide on how you want them organized,
# add tags so you can refer to sub plots in the text, etc. see below
p1 + p2 + plot_layout(ncol = 2) +
  plot_annotation(tag_levels = c("a"), tag_suffix = ")") &
  theme(
    plot.tag.position = c(0, 1),
    plot.tag = element_text(color = "darkred")
  )
```

```{r yourtablenamehere, echo=FALSE}
knitr::kable(mtcars[1:5, 1:5], caption = "Your table caption here")
```

You can include an equation using LaTeX code, 
for example see equation \@ref(eq:exampleequation).

\begin{equation}
    \theta^{*}|L^{*}, \delta^{2},Y,\sigma^{2}_{u} \sim N\left(B^{-1}W'L^{*},\delta^{2}B^{-1} \right)
    (\#eq:exampleequation)
\end{equation}

#### Fourth level heading (but don't use this many)
No need to deal with specifying superscript or subscript as you can just use 
the relevant unicode symbol directly in either the body text or in code chunks, 
e.g. ft², m³, CO₂, ½, °C.

<!-- You can copy paste in many more subscript/superscript symbols from here: 
https://en.wikipedia.org/wiki/Unicode_subscripts_and_superscripts and there 
are even more symbols here: https://en.wikipedia.org/wiki/Unicode_block. Also, this is how you add a comment in your body text. -->

# Aiding the review process
Though the output is a formatted Word file, the review process is still a bit 
clunky. I generally turn on track changes and send to co-authors for review. 
Then you manually bring those changes back into the .qmd file.  
Sometimes it's handy to add a 
[That way you can comment on why something was done a particular way, 
or propose an alternate, etc.,and folks can respond 
in that thread]{.comment-start id="0" author="The Flying Spahetti Monster" date="2300-01-01T10:12:00Z"}comment in the actual Word file output.[]{.comment-end id="0"}

# CRediT authorship contribution statement {.unnumbered}
**Author1Name**: Conceptualization, Data curation, Formal analysis,
Funding acquisition, Methodology, Investigation, Project administration,
Software, Supervision, Writing - original draft, Writing - review &
editing.

# Declaration of competing interest {.unnumbered}
All authors declare XYZ competing interest. The Center for the Built Environment (CBE) at the University of California, Berkeley, with which one authors is affiliated, is advised and funded in part by many approximately 50 partners that represent a diversity of organizations from the building industry, including manufacturers, building owners, facility managers, contractors, architects, engineers, government agencies, and utilities.

# Acknowledgements {.unnumbered}
The Funding Agency (Grant: XYZ) and the Center for the
Built Environment at the University of California, Berkeley supported
this research. Don't forget to thank any other folks who contributed!

# References {#references .unnumbered}
The string used to cite a reference is what's used to link the Bibtex file (references.bib) to a specific reference. You can use the BetterBibtex plugin within Zotero to automatically create these for your project, and to write the associated .bib file.