### 2024-10-18 word/blank.qmd
tidy to only required files

---
title: "TODOTitle"
author:
  - TODOAuthor1 ^[Center for the Built Environment, University of California Berkeley, USA] ^[Correspondence to TODO@berkeley.edu], TODOAuthor2AtSameOrg ^1^, TODOAuthor3atDifferentOrg ^[TODODifferentOrg]
date: "`r Sys.Date()`"
output:
  bookdown::word_document2:
    reference_docx: "./template.docx" 
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib 
csl : elsevier-with-titles.csl 
---

```{r knitrsetup, include = FALSE, cache = FALSE}
# knitr setup
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F,
  dev = "jpeg",
  cache = T,
  dpi = 300,
  fig.path = "figs",
  fig.show = "hold",
  fig.pos = "b",
  fig.width = 6.5, fig.height = 8.5, # these are the default figure sizes
  fig.path = "./figs"
)

# str(knitr::opts_chunk$get()) # run this in the console for all knitr config options
```

```{r librariesandthemesettings, include = FALSE, cache = FALSE}
# load all required libraries and configure themes

require(pacman)

# load packages using pacman
pacman::p_load(
  tidyverse, lubridate, here, rmarkdown, scales, patchwork, magrittr, janitor, # general, commonly used
  measurements, # most unit conversions 
  styler, # auto style your code with this addin
  # ggpubr, # ggplot extension focuses on annotations and fitted models
  # ggExtra, # ggplot extension handy for things like marginal density curves on figures
  # sjstats, pwr, # anova results
  # slider, # moving averages
  # rstatix, #pipe friendly stats
  palmerpenguins # quarto functionality
)

options(
  scipen = 999, # turn off scientific notation,
  digits = 2, # set default significant digits to 2, note this only works after a decimal point, use round() or format() to reduce precision of larger numbers
  timeout = 60
)

# set directory
here::i_am("blank.qmd")

# or theme light, with some  Cal colors
# figure theme setup
# set base font size for plots
theme_set(theme_light(base_size = 10))
# cleaner display of 'strips' above facets in plots
theme_update(
  strip.background = element_rect(fill = "#F5F5F5", color = "grey"),
  strip.text = element_text(color = "black"),
  strip.text.y = element_text(angle = 90)
)
#  lighten gridlines so just visible
theme_update(panel.grid = element_line(linetype = "dotted"))
# define a reasonably print and color-blind friendly color palette
# these also conveniently happen to be colors from the official UC Berkeley color palette. https://brand.berkeley.edu/identity-and-visual-design/color
cal_palette2 <- c("#3B7EA1", "#C4820E")
# with slight mod to dark yellow for better grayscale performance for 4 color scale
cal_palette4 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515")
# adding 5th color
cal_palette5 <- c("#003262", "#b0750c", "#3b7ea1", "#fdb515", "#D9661F")
```

```{r constants}
# user defined constants specific to this analysis
```

```{r functions}
# user defined functions specific to this analysis
```

```{r dataprep, include = FALSE}
# # Read in data, misc initial prep/cleaning tasks
# # for example, list all the files in a folder that match a specific pattern (e.g. csv files)
# 
# files <- list.files(path = "./input_file_folder/", pattern = "*.csv")
# 
# # create a short function to read in those files one by one
# # modify those files after you read them in,
# # and return a dataframe
# read_and_append_file <- function(fle) {
#   df <- read_csv(str_c("./input_file_folder/", fle))
#   # do whatever individual file level modifications you need here
#   # for example, extract some string from the filename
#   sss <- str_extract(string = fle, pattern = "(?<=df_).*(?=_USA_Clima)")
#   df %<>% mutate(sourcefilename = fle, stringofinterest = sss)
#   return(df)
# }
# 
# # apply that function and return an assembled dataframe
# df <- files %>% map_dfr(read_and_append_file)
# 
# # apply functions to tidy up the dataframe as necessary
# df %<>% clean_names()
```

# Abstract {.unnumbered}

# Introduction

## Literature review

# Methods

# Results

# Discussion

## Limitations

# CRediT authorship contribution statement {.unnumbered}
TODO **Author1Name**: Conceptualization, Data curation, Formal analysis,
Funding acquisition, Methodology, Investigation, Project administration,
Software, Supervision, Writing - original draft, Writing - review &
editing.

# Declaration of competing interest {.unnumbered}
TODO All authors declare XYZ competing interest. The Center for the Built Environment (CBE) at the University of California, Berkeley, with which one authors is affiliated, is advised and funded in part by many approximately 50 partners that represent a diversity of organizations from the building industry, including manufacturers, building owners, facility managers, contractors, architects, engineers, government agencies, and utilities.

# Acknowledgements {.unnumbered}
TODO The Funding Agency (Grant: XYZ) and the Center for the
Built Environment at the University of California, Berkeley supported
this research. Don't forget to thank any other folks who contributed!

# References {#references .unnumbered}